### 2019-10-24 paper/achievement_gap.Rmd
Starts adding plots to paper

---
title: Fake title
runninghead: Cimentada
author:
- name: Jorge Cimentada
  num: 1
address:
- num: 1
  org: Laboratory of Digital and Computational Demography, Max Planck Institute of Demographic Research
corrauth: "Konrad-Zuse-Str. 1. Rostock, Germany"
email: cimentada@demogr.mpg.de
abstract: "The literature on achievement inequality has recently started to focus on the dynamics of the socio-economic achievement gap in cognitive abilities. The main findings come from research in the U.S. revealing that the 90th/10th income achievement gap has widened 50% in the last 30 years. This chapter aims to investigate whether there are patterns in the evolution of the achievement gap from a comparative perspective. Using 15 years of data in 32 countries from the Program for International Student Assessment (PISA), I find that there is considerable variation in the way in which the gap between the average score of students above (and at) the 90th percentile and below (and) the 10th percentile is evolving. The prime examples come from the U.S. and Germany closing at about 50% and 30% in the last 15 years while France is widening at a similar rate. I find that curricular tracking and vocational enrollment explain 40% of the variance in the achievement gap between countries and show that the relationship is conditioned by a strong interaction. Low curricular tracking is associated with a small achievement gap, whereas high levels of curricular tracking is associated with wide achievement gaps. However, once tracking is coupled with high vocational enrollment this can remedy the potential adverse effects and reduce the gap by .6 standard deviation. I use simulations to show that switching to less curricular tracking can help decrease a country’s SES gap by about 10% while switching to more tracking would increase the achievement gap by about 51% percent."
keywords: academic achievement, education inequality, school autonomy, international comparison
classoption:
  - Royal
  - times
bibliography: bibfile
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
---

```{r }
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  ## fig.width = 6,
  ## fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  progress = FALSE,
  verbose = FALSE,
  results = "asis"
)
```

# The Body of the Article

# Bibliography

Link a `.bib` document via the YAML header, and bibliography will be printed at the very end (as usual). The default bibliography style is provided by Wiley as in `WileyNJD-AMA.bst`, do not delete that file.

Use the Rmarkdown equivalent of the \LaTeX citation system using `[@<name>]`. Example: [@Taylor1937], [@Knupp1999; @Kamm2000].

To include all citation from the `.bib` file, add `\nocite{*}` before the end of the document.

```{r, fig.height = 90, fig.width = 60, out.width = "95%"}
p1_plot <- readd(p1_evolution_gaps)
p1_plot
```

# Further information

All \LaTeX enviroments supported by the main template are supported here as well; see the `.tex` sample file [here](http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1097-0258/homepage/la_tex_class_file.htm) for more details and example.